\textsc{Atomic modeling of argon}\\ \footnotesize{\textsc{Project 5 --- FYS3150}}

### 2015-11-27 projects/project_5/molecular_dynamics/article/article.tex
fixed the methods

\documentclass{article}
\usepackage[]{enumerate} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{amsthm, amsmath, amssymb} 
\usepackage[]{hyperref} 
\graphicspath{{../pictures/}}

@@

    Argon has a melting point of $\approx 84$K and a boiling point of $\approx
    87$K. If we run the simulations, once for a temperature of $40$K, once for
    $85$K and finally once for $120K$ we should see indications of three
    different phases. The results shows that the model somewhat accurately
    manages to capture these phase transitions. Snapshots taken at the same
    point in time for all three cases are presented in figure
    \ref{fig:potential_iteration}.
    \begin{figure}
        \begin{tabular}[c]{ccc}
            \begin{subfigure}[c]{0.32\textwidth}
                \includegraphics[height=100pt, width=\linewidth]{solid_40K.png}
                \caption{$T = 40$K}
            \end{subfigure}
            \begin{subfigure}[c]{0.32\textwidth}
                \includegraphics[height=100pt, width=\linewidth]{liquid_85K.png}
                \caption{$T = 85$K}
            \end{subfigure}
            \begin{subfigure}[c]{0.32\textwidth}
                \includegraphics[height=100pt, width=\linewidth]{gas_120K.png}
                \caption{$T = 120$K}
            \end{subfigure}
        \end{tabular}
        \caption[Argon phases]{Snapshots taken of the system at frame 500/999 for
            temperatures of $40$, $85$ and $120$ kelvin. It is a bit difficult
            to conclude from the pictures, one really needs to see the system
            in motion, but the system do behave more or less like a solid, a
            liquid and a gas for respective temperatures. Any artifacts can be
            contributed to numerical instability. It is possible to see a
            certain change in the entropy of the system across the three
            images. These made use of the Euler-Cromer integration method.}
        \label{fig:potential_iteration}
    \end{figure}

@@

    We are also interested in the instantaneous temperature of the system, and
    the equipartition theorem \cite{equipar} can help us with this. The theorem
    relates the average energy of a system to its temperaturkke and gives us the
    relation $\langle E_k \rangle = (3/2)k_bT$ which can be solved for
    temperature:
    \begin{equation}
        \notag
        T = \frac{2}{3} \frac{E_k}{nk_b}.
    \end{equation}
    Notice the removal of the brackets --- we are only interested in the
    \emph{instantaneous} temperature at any point in time. We also note that
    since our units are scaled, Boltzmann's constant is equal to one.
    The density of the system is the total mass of the system divided by the volume.

>>>>

\documentclass{article}
\usepackage[]{enumerate} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{amsthm, amsmath, amssymb} 
\usepackage[]{hyperref} 
\urlstyle{sf}
\graphicspath{{../pictures/}}

@@

    We need to chose the parameters $\sigma$ and $\varepsilon$ so that they are
    suited for simulation of argon. It turns out that $\sigma = 3.405
    \mathrm{\AA}$ and $\varepsilon = 119.8k_B\mathrm{K}$ with $k_B = 1$ in
    MD-units.

    Argon has a melting point of $\approx 84$K and a boiling point of $\approx
    87$K. If we run the simulations, once for a temperature of $40$K, once for
    $85$K and finally once for $120K$ we should see indications of three
    different phases. The results shows that the model somewhat accurately
    manages to capture these phase transitions. Snapshots taken at the same
    point in time for all three cases are presented in figure
    \ref{fig:potential_iteration}.
    \begin{figure}
        \begin{tabular}[c]{ccc}
            \begin{subfigure}[c]{0.32\textwidth}
                \includegraphics[height=100pt, width=\linewidth]{solid_40K.png}
                \caption{$T = 40$K}
            \end{subfigure}
            \begin{subfigure}[c]{0.32\textwidth}
                \includegraphics[height=100pt, width=\linewidth]{liquid_85K.png}
                \caption{$T = 85$K}
            \end{subfigure}
            \begin{subfigure}[c]{0.32\textwidth}
                \includegraphics[height=100pt, width=\linewidth]{gas_120K.png}
                \caption{$T = 120$K}
            \end{subfigure}
        \end{tabular}
        \caption[Argon phases]{Snapshots taken of the system at frame 500/999 for
            temperatures of $40$, $85$ and $120$ kelvin. It is a bit difficult
            to conclude from the pictures, one really needs to see the system
            in motion, but the system do behave more or less like a solid, a
            liquid and a gas for respective temperatures. It is possible to see a
            certain change in the entropy of the system across the three
            images. These made use of the Velocity-Verlet integration method.}
        \label{fig:potential_iteration}
    \end{figure}

@@

    We are also interested in the instantaneous temperature of the system, and
    the equipartition theorem \cite{equipar} can help us with this. The theorem
    relates the average energy of a system to its temperaturkke and gives us the
    relation $\langle E_k \rangle = (3/2)k_bT$ which can be solved for
    temperature:
    \begin{equation}
        \notag
        T = \frac{2}{3} \frac{E_k}{nk_b}.
    \end{equation}
    Notice the removal of the brackets --- we are only interested in the
    \emph{instantaneous} temperature at any point in time. We also recall that
    since our units are scaled, Boltzmann's constant is equal to one.

In this section we consider various systems and simulate them using the program
devised in the previous section.

\subsection{Energy conservation of integration method}
\label{sub:energy_conservation_of_integration_method}

We initially decided to use the Velocity-Verlet integration method as opposed
to Euler-Cromer for its good energy conservation. We now simulate a set of
similar systems with varying temperature and look at how the fluctuations in
total energy scales with the timestep $\Delta T$. In order to do this we can
consider the standard-devation of the total energy.

\subsection{Melting temperature of system}
\label{sub:melting_temperature_of_system}
